---
title: "HW2"
author: "Vivian Xia"
date: "2024-09-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1

```{r}
# Load necessary libraries
library(dplyr)

# Read the CSV file
data <- read.csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

# Select relevant columns
cleaned_data <- data %>%
  select("Line","Station.Name", "Station.Latitude", "Station.Longitude", c("Route1":"Route11"),"Entry","Vending", "Entrance.Type","ADA") %>%
  
  # Convert Entry to logical (TRUE for "YES", FALSE for "NO")
  mutate(Entry = ifelse(Entry == "YES", TRUE, FALSE))

# View the data
View(cleaned_data)
```

The dataset provides details about NYC subway entrances and exits. Key variables include:Line (Subway line), Station Name(Name of the subway station), Station Latitude/Longitude(Geographical coordinates of the station), Routes Served (Route1 to Route11), Entry(Logical variable indicating whether entry is allowed at a specific entrance), Vending(Indicates whether there are vending machines at the entrance),Entrance Type(e.g. staircase, elevator), ADA(Whether the entrance is ADA-compliant)

So far, I have selected relevant columns from the original data set and converted the Entry variable into a logical format. The resulting dataset has 1868 rows and 19 columns. This dataset can be considered tidy, as each variable has its own column, each observation has its own row, and each value has its own cell.

```{r}
# Count distinct stations based on the combination of station name and line
distinct_stations <- cleaned_data %>%
  distinct(Line,Station.Name) %>%
  nrow()

distinct_stations
```
There are 465 stations.

```{r}
# Count ADA compliant stations
ada_compliant_stations <- cleaned_data %>%
  filter(ADA == TRUE) %>%
  nrow()

ada_compliant_stations
```
There are 468 stations.

```{r}
# Calculate the proportion of entrances without vending that allow entry
proportion_without_vending <- cleaned_data %>%
  filter(Vending == "NO") %>%
  summarise(proportion = mean(Entry))

proportion_without_vending
```
The proportion of entrances without vending is approximately 37.7%.

```{r}
library(tidyr)
   cleaned_data <- cleaned_data %>%
     mutate(across(Route1:Route11, as.character))
   
# Transfer data to long character format
routes_data <- cleaned_data %>%
  pivot_longer(Route1:Route11, names_to = "RouteNumber", values_to = "Route")
routes_data <- routes_data %>%
  filter(!is.na(Route))

# Check the reformatted data
View(routes_data)
```

```{r}
# Count distinct stations serving the A train
a_train_stations <- routes_data %>%
  filter(Route == "A") %>%
  distinct(Line, Station.Name) %>%
  nrow()

a_train_stations
```
There are 60 distinct stations serving the A train,

```{r}
# Count ADA compliant stations serving the A train
ada_compliant_a_train <- routes_data %>%
  filter(Route == "A", ADA == TRUE) %>%
  distinct(Line, Station.Name) %>%
  nrow()

ada_compliant_a_train
```
There 17 ADA compliant stations serving the A train.


Problem 2

```{r}
# Load necessary libraries
library(readxl)

# Read Mr. Trash Wheel sheet
mr_trash_wheel <- read_excel("202409 Trash Wheel Collection Data 2.xlsx",sheet = "Mr. Trash Wheel", skip = 1)

# Clean the data by selecting relevant columns and renaming them
mr_trash_wheel_cleaned <- mr_trash_wheel %>%
 select(Dumpster, Month, Year, Date, 'Weight.tons'= 'Weight (tons)', 'Volume.cubicyards' =     'Volume (cubic yards)', PlasticBottles ='Plastic Bottles' , Polystyrene, CigaretteButts = 'Cigarette Butts', GlassBottles = 'Glass Bottles', PlasticBags = 'Plastic Bags', Wrappers, SportsBalls = 'Sports Balls' ,HomesPowered = 'Homes Powered*', fifteen ='...15', sixteen='...16') %>%
  
  # omit rows that do not include dumpster-specific data
  filter(!is.na(Dumpster)) %>%
  # round the number of sports balls to the nearest integer and converts the result to an    integer variable
  mutate(SportsBalls = as.integer(round(SportsBalls))) %>%
  
  # add an additional variable to keep track of Trash Wheel
  mutate(Trash_Wheel = "Mr. Trash Wheel") %>%
  # transfer year to numerical character
  mutate(Year = as.numeric(Year))

View(mr_trash_wheel_cleaned)
```


```{r}
# Read Professor Trash Wheel sheet
prof_trash_wheel <- read_excel("202409 Trash Wheel Collection Data 2.xlsx", sheet = "Professor Trash Wheel", skip = 1)

# Clean
prof_trash_wheel_cleaned <- prof_trash_wheel %>%
  select(Dumpster, Month, Year, Date, 'Weight.tons'= 'Weight (tons)', 'Volume.cubicyards' =     'Volume (cubic yards)', PlasticBottles ='Plastic Bottles' , Polystyrene, CigaretteButts = 'Cigarette Butts', GlassBottles = 'Glass Bottles', PlasticBags = 'Plastic Bags', Wrappers, HomesPowered = 'Homes Powered*') %>%
  
# omit rows that do not include dumpster-specific data
  filter(!is.na(Dumpster)) %>%
  
# add an additional variable to keep track of Trash Wheel
  mutate(Trash_Wheel = "Professor Trash Wheel")

# Repeat for Gwynnda
gwynnda <- read_excel("202409 Trash Wheel Collection Data 2.xlsx", sheet = "Gwynnda Trash Wheel", skip = 1)

gwynnda_cleaned <- gwynnda %>%
  select(Dumpster, Month, Year, Date, 'Weight.tons'= 'Weight (tons)', 'Volume.cubicyards' =     'Volume (cubic yards)', PlasticBottles ='Plastic Bottles' , Polystyrene, CigaretteButts = 'Cigarette Butts', PlasticBags = 'Plastic Bags', Wrappers, HomesPowered = 'Homes Powered*') %>%
  
filter(!is.na(Dumpster)) %>%

# add an additional variable to keep track of Trash Wheel
  mutate(Trash_Wheel = "Gwynnda Trash Wheel")

# Combine all data set
all_trash_wheel_data <- bind_rows(mr_trash_wheel_cleaned,
                                  prof_trash_wheel_cleaned,
                                  gwynnda_cleaned)

# Check the combined data set
View(all_trash_wheel_data)
```

The `all_trash_wheel_data` combines the data from three different trash wheels of Healthy Harbor company from the year 2014 to 2024, if data is available. The data includes the number of Dumpster, Month, Year, Date, weight, volumes, different categories of trash collected, and homes powered. 

```{r}
# Total weight collected by Professor Trash Wheel
total_weight_professor <- all_trash_wheel_data %>%
  filter(Trash_Wheel == "Professor Trash Wheel") %>%
  summarise(total_weight = sum(Weight.tons, na.rm = TRUE)) %>%
print()
```
The total weight collected by Professor Trash Wheel is 246.74 tons.

```{r}
# Cigarette butts collected by Gwynnda in June 2022
gwynnda_cigarette_june <- all_trash_wheel_data %>%
  filter(Trash_Wheel == "Gwynnda Trash Wheel", Month == "June", Year == 2022) %>%
  summarise(total_cig_butts = sum(CigaretteButts, na.rm = TRUE))%>%
print()
```
The cigarette butts collected by Gwynnda in June 2022 is 18120.



Problem 3

```{r}
# Load necessary libraries
library(readr)

# Read the CSV files
bakers_df <- read_csv("bakers.csv")
bakes_df <- read_csv("bakes.csv")
results_df <- read_csv("results.csv")

# make the data tidy
results_df <- results_df %>%
  slice(3:n()) %>% #keep the data from the third row
  setNames(c("Series", "Episode", "Baker", "Technical", "Result")) 

# make the data consistent
bakes_df <- bakes_df %>%
  mutate(Series = as.numeric(Series), Episode = as.numeric(Episode))
  
results_df <- results_df %>%
  mutate(Series = as.numeric(Series), Episode = as.numeric(Episode))

bakers_df <- bakers_df %>%
  rename(Baker = `Baker Name`) %>%
  mutate(Series = as.numeric(Series)) %>%
  mutate(Baker = ifelse(Baker == 'Joanne', 'Jo',Baker))


View(bakers_df)
View(bakes_df)
View(results_df)
```


```{r}
# Check the completeness and correctness
bakers_df <- bakers_df %>%
  mutate(Baker = word(Baker, 1))
```

```{r}
# Check the not matching one

resultbakernumber<- anti_join(results_df, bakers_df, by = c("Series", "Baker")) |>
  nrow()

print(resultbakernumber)

bakerresultnumber <- anti_join(bakers_df, results_df, by = c("Series", "Baker")) |>
  nrow()

print(bakerresultnumber)
```

```{r}
# Merge the bakers and the results


```

